---
title: "p8105_hw3_xs2447"
output: github_document
---
Basic setting
```{r}
library(tidyverse)
library(leaflet)
library(gghighlight)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

##Problem 1

load in Instacart dataset

```{r}
library(p8105.datasets)
data("instacart")
```

There are 1384617 rows and 15 variables. Each row of the dataset is a product from an order and the product from the same product_id are listed in adjacent rows. Some important variables of the dataset are order id, product_id, used_id, aisle_id and department_id, which are identifiers. Besides, order_dow is also an important variable to indicated the day of the week on which the order was placed. For example, order number 1 made by user 112108 has 8 products in it and it is placed 10am on Thursday. 4 of the 8 products belong to produce department, while the rest come from dairy eggs and canned goods department. 


*How many aisles are there, and which aisles are the most items ordered from?

```{r}
aisle_df = instacart %>% 
  group_by(aisle,aisle_id) %>% 
  summarize(n_obs = n()) %>% 
  arrange(desc(n_obs))

aisle_df
```

There are 134 aisles and fresh vegetables has the most (150609 in total) ordered items.


*Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.

```{r}
aisle_df %>% 
  filter(n_obs > 10000) %>% 
  ggplot(aes(x = aisle, y = n_obs)) +
  geom_col() +
  coord_flip() 
```

There are 39 aisles with more than 10000 items ordered. From these 39 aisles, fresh vegetables aisle has the most items, while butter has the least items.


*Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r}
three_pop = instacart %>% 
  filter(aisle %in% c("baking ingredients","dog food care", "packaged vegetables fruits")) %>% 
  group_by(aisle) %>% 
  count(product_name, name = "product_count") %>% 
  mutate(product_rank = min_rank(desc(product_count))) %>% 
  filter(product_rank <= 3) %>% 
  arrange(aisle, product_rank) %>% 
  knitr::kable()

three_pop
```

The most popular products in baking ingredients are Light Brown Sugar, Pure Baking Soda and Cane Sugar. From dog food care, the most three popular products are Snack Sticks Chicken & Rice Recipe Dog Treats, Organix Chicken & Brown Rice Recipe, and Small Dog Biscuits. From packaged vegetables fruits, Organic Baby Spinach, Organic Raspberries, and Organic Blueberries are most popular.


*Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

```{r}
mean_hour_of_day = instacart %>% 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  group_by(order_dow) %>% 
  summarize(mean_hour = mean(order_hour_of_day)) %>% 
  knitr::kable()

mean_hour_of_day
```

For Pink Lady Apples and Coffee Ice Cream, the mean hour of order is at mid of the day on each day, the mean ranges from 12.17391 to 14.68519.

## Problem 2

load BRFSS dataset

```{r}
data("brfss_smart2010")
```

*format the data to use appropriate variable names
*focus on the “Overall Health” topic
*include only responses from “Excellent” to “Poor”
*organize responses as a factor taking levels ordered from “Poor” to “Excellent”

```{r}
brfss = brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  filter(topic == "Overall Health") %>%
  filter(response %in% c("Poor","Fair", "Good", "Very good", "Excellent")) %>% 
  arrange(factor(response, c("Poor","Fair", "Good", "Very good", "Excellent"))) %>% 
  rename(state = locationabbr, location = locationdesc)
```


*In 2002, which states were observed at 7 or more locations? What about in 2010?

```{r}
brfss_2002 = brfss %>% 
  filter(year == 2002) %>% 
  group_by(state) %>% 
  summarize(obs_7 = n_distinct(location)) %>% 
  filter(obs_7 >= 7)
  
brfss_2002
```

There are 6 states were observed at 7 or more locations in 2002, including CT, FL, MA, NC, NJ and PA.

```{r}
brfss_2010 = brfss %>% 
  filter(year == 2010) %>% 
  group_by(state) %>% 
  summarize(obs_7 = n_distinct(location)) %>% 
  filter(obs_7 >= 7)
  
brfss_2010
```

There are 14 states were observed at 7 or more locations in 2010, including CA, CO, FL, MA, MD, NC, NE, NJ, NY, OH, PA, SC, TX, and WA.


*Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).

```{r}
brfss %>% 
  filter(response == "Excellent") %>% 
  group_by(state, year) %>% 
  summarize(means = mean(data_value)) %>% 
  ggplot(aes(x = year, y = means, color = state)) +
  geom_line(alpha = .6) 

```

From the plot, we can find that maximum means of the data_value is generally decreasing from 2002 to 2010, while the minimum means flutuate a lots.

```{r}
brfss %>% 
  filter(response == "Excellent") %>% 
  group_by(state, year) %>% 
  summarize(means = mean(data_value)) %>% 
  ggplot(aes(x = year, y = means, color = state)) +
  geom_line(alpha = .6) +
  gghighlight(state == "WV") 
```

If we highlight the WV state, we can also find that for most of the time, the mean value of data in this state is lower than others'. 


*Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.
```{r}
brfss %>% 
  filter(state == "NY", year == c(2006, 2010)) %>% 
  ggplot(aes(x = response)) +
  geom_bar() + 
  facet_grid(.~year)
```

From the plot, we can see that the responses is equally distributed in 2006 while in 2010, more people choose good or poor. We can also find we get more response in 2010 compare to 2006. 

## Problem 3
*Load, tidy, and otherwise wrangle the data. Your final dataset should include all originally observed variables and values; have useful variable names; include a weekday vs weekend variable; and encode data with reasonable variable classes. Describe the resulting dataset (e.g. what variables exist, how many observations, etc).
```{r}
accel_df = read_csv("accel_data.csv")
```

```{r}
accel_tidy_df = accel_df %>% 
  mutate(
    week_doe = if_else(day %in% c("Saturday", "Sunday"), "weekend","weekday")
    ) %>% 
  relocate(week, day_id, day, week_doe) %>% 
  pivot_longer(activity.1:activity.1440,
               names_to = "minute",
               names_prefix = "activity.",
               values_to = "activity_count") %>% 
  mutate(minute = as.double(minute))
```

The original dataset has 35 rows and 1143 columns. After tidying up, the dataset has 50400 rows and 6 variables, including the week number, the day number, the day of the week, weekday or weekend, minute number and the activity count at each specific time. The dataset records the activity count over 35 days.


*Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate accross minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent?

```{r}
accel_tidy_df %>% 
  group_by(day_id) %>% 
  summarize(day_total = sum(activity_count)) %>% 
  knitr::kable()
```

The total counts of activity fluctuate significantly, range from 1440 to 685910, so there is no apparent trend. 

*Accelerometer data allows the inspection activity over the course of the day. Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph.
```{r}
accel_tidy_df %>% 
  ggplot(aes(x = minute, y = activity_count, color = day)) +
  geom_point(alpha = .3) +
  geom_smooth(alpha = .7, se = FALSE) + 
  scale_x_continuous(breaks = seq(0, 1440, 120))
```

From the plot, we can see that the activity count is relatively low from about 0 to 300 minute, and from 1320 to 1440 minute. These minutes are about 10pm to 6am. Therefore, generally speaking, the activity count is low when people are asleep. The peaks of the activity count often happen from about 360 to 720 minute (6am to 12pm), 960 to 1020 (4pm to 5pm), and 1140 to 1320 (7pm to 10pm). At these times, people are more likely doing activity, therefore the count is high. Besides, we can also find that the people usually doing more activity at Friday night. On Wednesday, this people are usually doing heaviest activity at about 7:30pm. 

