---
title: "p8105_hw3_xs2447"
output: github_document
---
Basic setting
```{r}
library(tidyverse)
library(leaflet)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

##Problem 1

load in Instacart dataset

```{r}
library(p8105.datasets)
data("instacart")
```

There are 1384617 rows and 15 variables. Each row of the dataset is a product from an order and the product from the same product_id are listed in adjacent rows. Some important variables of the dataset are order id, product_id, used_id, aisle_id and department_id, which are identifiers. Besides, order_dow is also an important variable to indicated the day of the week on which the order was placed. For example, order number 1 made by user 112108 has 8 products in it and it is placed 10am on Thursday. 4 of the 8 products belong to produce department, while the rest come from dairy eggs and canned goods department. 


*How many aisles are there, and which aisles are the most items ordered from?

```{r}
aisle_df = instacart %>% 
  group_by(aisle,aisle_id) %>% 
  summarize(n_obs = n()) %>% 
  arrange(desc(n_obs))

aisle_df
```
There are 134 aisles and fresh vegetables has the most (150609 in total) ordered items.


*Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.

```{r}
aisle_df %>% 
  filter(n_obs > 10000) %>% 
  ggplot(aes(x = aisle, y = n_obs)) +
  geom_col() +
  coord_flip() 
```
There are 39 aisles with more than 10000 items ordered. From these 39 aisles, fresh vegetables aisle has the most items, while butter has the least items.


*Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r}
three_pop = instacart %>% 
  filter(aisle %in% c("baking ingredients","dog food care", "packaged vegetables fruits")) %>% 
  group_by(aisle) %>% 
  count(product_name, name = "product_count") %>% 
  mutate(product_rank = min_rank(desc(product_count))) %>% 
  filter(product_rank <= 3) %>% 
  arrange(aisle, product_rank) %>% 
  knitr::kable()

three_pop
```
The most popular products in baking ingredients are Light Brown Sugar, Pure Baking Soda and Cane Sugar. From dog food care, the most three popular products are Snack Sticks Chicken & Rice Recipe Dog Treats, Organix Chicken & Brown Rice Recipe, and Small Dog Biscuits. From packaged vegetables fruits, Organic Baby Spinach, Organic Raspberries, and Organic Blueberries are most popular.


*Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

```{r}
mean_hour_of_day = instacart %>% 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  group_by(order_dow) %>% 
  summarize(mean_hour = mean(order_hour_of_day)) %>% 
  knitr::kable()

mean_hour_of_day
```
For Pink Lady Apples and Coffee Ice Cream, the mean hour of order is at mid of the day on each day, the mean ranges from 12.17391 to 14.68519.

## Problem 2

load BRFSS dataset

```{r}
data("brfss_smart2010")
```

*format the data to use appropriate variable names
*focus on the “Overall Health” topic
*include only responses from “Excellent” to “Poor”
*organize responses as a factor taking levels ordered from “Poor” to “Excellent”

```{r}
brfss = brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  filter(topic == "Overall Health") %>%
  filter(response %in% c("Poor","Fair", "Good", "Very good", "Excellent")) %>% 
  arrange(factor(response, c("Poor","Fair", "Good", "Very good", "Excellent")))
```


*In 2002, which states were observed at 7 or more locations? What about in 2010?


*Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).


*Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.